---
title: "MeganticTraits_Species~Traits" 
author: "Julie Messier" 
output: 
   pdf_document:
      toc: true
      toc_depth: 3
      fig_width: 4.5
      fig_height: 3.5
date: '`r Sys.Date()`'
---
***


```{r,include=FALSE}
#<<LIBRARIES>>
library(car) # for powerTransform
library(vegan) # for decostand (standardizing data)
library(tree) # for tree()
library(MASS) # for glm.nb()  
library(mgcv) # for gam()
library(MuMIn)

wrk.dir<-("C:/Users/Julie/Desktop/Postdoc/PROJECT - Megantic Traits/Workspaces/") # Workspaces
data.dir<-(("C:/Users/Julie/Desktop/Postdoc/PROJECT - Megantic Trait/Data/")) # data
res.dir<-("C:/Users/Julie/Desktop/Postdoc/PROJECT - Megantic Traits/Results/")  # Results
grp.dir<-("C:/Users/Julie/Desktop/Postdoc/PROJECT - Megantic Trait/Graphs/")   # Graphs
fct.dir<-("C:/Users/Julie/Desktop/Postdoc/PROJECT - Megantic Trait/Functions/") # Functions

```    

```{r,echo=F,eval=T, include=FALSE}
#<<LOAD FILES>>

#U.traits4.sp
load(file=paste0(wrk.dir,'Species-Level.Traits.Understory-Layer.with.MycFrac.and.SeedSize.and.PCs.RData')) 
dim(U.traits4.sp)

# U.Ys.c
load(file=paste0(wrk.dir,'All.Response.variables.Understory.Layer.Inf.removed.RData')) 
dim(U.Ys.c)

# subset abundance data to include only the species with trait data 
    sp.response<-U.Ys.c[rownames(U.Ys.c)%in%rownames(U.traits4.sp),]
    dim(sp.response)
    
# subset trait data to include only the species with abundance data
Xs<-U.traits4.sp[rownames(U.traits4.sp)%in%rownames(U.Ys.c),]
dim(Xs)

# Create long and short lists of traits 
    Trait.Names.8t<-c("Log.Ht.veg","Max.Root.Loca","Log.Lamina.thck","LDMC","Log.Leaf.Area",  
                        "Leaf.Mass.Frac","SRL","Myc.Frac")
    length(which(complete.cases(Xs[,Trait.Names.8t])))
      # 36 species
    
    Trait.Names.5t<-c("Log.Ht.veg","Log.Lamina.thck","LDMC","Log.Leaf.Area","Leaf.Mass.Frac")
    length(which(complete.cases(Xs[,Trait.Names.5t])))
      # 46 species
    
Xs<-U.traits4.sp[rownames(U.traits4.sp)%in%rownames(U.Ys.c),]

    dat.8t<-merge(sp.response[,c("LatinName","abundance.ratio","occurence.ratio",'ElevDif')],
                    Xs[which(complete.cases(Xs[,Trait.Names.8t])),Trait.Names.8t],
                    by="row.names",all.y=T)
    # Cleanup
    dim(dat.8t)
    # 36 13
    rownames(dat.8t)<-dat.8t$Row.names
    dat.8t$Row.names<-NULL
    dim(dat.8t)
    # 36 12
```

#(A) Abundance Ratio ~ 8 Traits

## A1 - Explore Trait Interactions
   
### A1.1 - Regression Tree
   
```{r,echo=F,eval=T}
   
    par(mfrow=c(1,1))
    tree.model.abund<-tree(sp.response$abundance.ratio~.,data=Xs[,Trait.Names.8t],
                     control=tree.control(nobs=36,mincut=3))
    plot(tree.model.abund,mar=c(3,1,5,1)); text(tree.model.abund)
    title('Abundance.ratio ~ 8 traits \n Deviance explained = 0.51')
      # Myc.frac is most important variable       # for sp. with high myc.frac, SRL matters
    tree.model.abund
```    

**Deviance explained:**
```{r,echo=T}
  1-(deviance(tree.model.abund)/178.1) 
```

>            =======
>            A1 - SUMMARY
>              trait-trait interactions to include in global model: 
>            
>                * Myc.Frac:SRL 
>                * SRL:Max.Root.Loca
>                * SRL:Log.Ht.veg
>                * Myc.Frac:SRL:Max.Root.Loca
>                * Myc.Frac:SRL:Log.Ht.veg
>                
>            =======




## A2 - Test Full Model
```{r,echo=F}
H.full.abund<-glm(abundance.ratio~
                    Myc.Frac+SRL+Max.Root.Loca+Log.Ht.veg+
                     Log.Lamina.thck+LDMC+Log.Leaf.Area+Leaf.Mass.Frac,
                 data=dat.8t,
                 family=Gamma(link='log'),
                 maxit=1000)
summary(H.full.abund)$coefficients
```
**Deviance explained:**
```{r,echo=T}
1-(H.full.abund$deviance/H.full.abund$null) 
AICc(H.full.abund)

# Analysis of Deviance to get p-value of whole model
H0<-glm(abundance.ratio~1,
+         data=dat.8t,
+         family=Gamma(link='log'),
+         maxit=1000)

with(anova(H0,H.full.abund),pchisq(Deviance,Df,lower.tail=FALSE)[2])
```

>         =======
>         A2 - SUMMARY
>         Only Mycorrhizal fraction is significant. 
>         Full model explains 36% of deviance with p=0.09
>         =======



## A3 - Find best subset of models with AICc dredge ####

```{r,echo=F}

   # Not including both Myc.Frac and log(Myc.Frac). Include:
   #        - all 1st order terms 
   #        - 2-way and 3-way trait interactions from regression tree
   #        - Don't include trait interaction from interaction screening.
         
          # rewrite H1 with na.action='fail' - required by dredge ()
          #
          
H.global.abund<-glm(abundance.ratio~
                        Myc.Frac+SRL+Max.Root.Loca+Log.Ht.veg+Log.Lamina.thck+LDMC+Log.Leaf.Area+Leaf.Mass.Frac+
                        Myc.Frac:SRL + SRL:Max.Root.Loca + SRL:Log.Ht.veg+
                        Myc.Frac:SRL:Log.Ht.veg + Myc.Frac:SRL:Max.Root.Loca,
                     data=dat.8t,
                     family=Gamma(link='log'),
                     maxit=1000,na.action='na.fail')
```
```{r,echo=T,eval=T,message=F}
dredge.abund<-dredge(H.global.abund,beta='sd',rank="AICc",m.lim = c(0,6),extra=c("R^2"),trace=F)
dredge.abund[1:5]

```

>         =======
>         A3- SUMMARY
>         Only Mycorrhizal fraction matters. 
>         It explains ~1/3 of the variation in change in abundance
>         Best models have 2 traits
>         =======

***
\pagebreak

#(B) Occurence Ratio ~ 8 Traits
- occurence and abundance are correlated. Do we want to present them both?

## B1 - Explore Trait Interactions
   
### B1.1 - Regression Tree

```{r,echo=F,eval=T}
par(mfrow=c(1,1))
     tree.model.occur<-tree(dat.8t$occurence.ratio~.,data=dat.8t[,Trait.Names.8t])
     plot(tree.model.occur,mar=c(3,1,5,1)); text(tree.model.occur);
     title('Occurence.ratio vs 8 traits \n Deviance explained =0.32')
     
     tree.model.occur
```
**Deviance explained:**
```{r,echo=T}
  1-(deviance(tree.model.occur)/30.79) 
```

>            =======
>            B1 - SUMMARY
.              trait-trait interactions to include in global model: 
>            
>                * Myc.Frac:SRL 
>                * SRL:Max.Root.Loca
>                * Max.Root.Loca:Log.Lamina.thck
>                * Myc.Frac:SRL:Max.Root.Loca
>                * SRL:Myc.Frac:Log.Lamina.thck
>                
>            =======

## B2 - Test Full Model
```{r,echo=F}
H.full.occur<-glm(occurence.ratio~
                    Myc.Frac+SRL+Max.Root.Loca+Log.Lamina.thck+Log.Ht.veg+
                     LDMC+Log.Leaf.Area+Leaf.Mass.Frac,
                 data=dat.8t,
                 family=Gamma(link='log'),
                 maxit=1000)
summary(H.full.occur)$coefficients
AICc(H.full.occur)
```
**Deviance explained:**
```{r,echo=T}
1-(H.full.occur$deviance/H.full.occur$null) 

# Analysis of Deviance to get p-value of whole model
H0<-glm(occurence.ratio~1,
+         data=dat.8t,
+         family=Gamma(link='log'),
+         maxit=1000)

with(anova(H0,H.full.occur),pchisq(Deviance,Df,lower.tail=FALSE)[2])
```

>         =======
>         B2 - SUMMARY
>         Nothing significant. 
>         Full model explains 36% of deviance with p=0.95 
>            - although no variables are significant...
>         =======

## B3 - Find best subset of models with AICc dredge ####

```{r,echo=F}


   # Not including both Myc.Frac and log(Myc.Frac). Include:
   #        - all 1st order terms 
   #        - 2-way and 3-way trait interactions from regression tree
   #        - Don't include trait interaction from interaction screening.
         
          # rewrite H1 with na.action='fail' - required by dredge ()
          #
          
H.global.occur<-glm(occurence.ratio~
                 Myc.Frac+SRL+Max.Root.Loca+Log.Lamina.thck+Log.Ht.veg+LDMC+Log.Leaf.Area+Leaf.Mass.Frac+
                 Myc.Frac:SRL+SRL:Max.Root.Loca+Max.Root.Loca:Log.Lamina.thck+
                 Myc.Frac:SRL:Max.Root.Loca,
              data=dat.8t,
              family=Gamma(link='log'),
              maxit=1000,na.action='na.fail')

```
```{r,echo=T,eval=T,message=F}
dredge.occur<-dredge(H.global.occur,beta='sd',rank="AICc",m.lim = c(0,6),extra=c("R^2"),trace=F)
dredge.occur[1:7]

```

>         =======
>         B3 - SUMMARY
>         Log.Leaf.Area matters. 
>         Traits explains ~1/5 of variation in change in occurence 
>            - although individually no variables are significant...
>         Best models have 1-3 traits
>         =======

#(C) Elevation Shift ~ 8 Traits

## C1 - Explore Trait Interactions
   
### C1.1 - Regression Tree

```{r,echo=F,eval=T}
par(mfrow=c(1,1))
     tree.model.elev<-tree(dat.8t$ElevDif~.,data=dat.8t[,Trait.Names.8t])
     plot(tree.model.elev,mar=c(3,1,5,1)); text(tree.model.elev);
     title('Elevation Shift ~ 8 traits \n Deviance explained = 0.60')
     
     tree.model.elev
```
**Deviance explained:**
```{r,echo=T}
  1-(deviance(tree.model.elev)/92210) 
```

***
\pagebreak

>            =======
>            C1 - SUMMARY
               trait-trait interactions to include in global model: 
>            
>                * Leaf.Mass.Frac:Max.Root.Loca
>                * Max.Root.Loca:Log.Ht.veg
>                * Leaf.Mass.Frac:Max.Root.Loca:Log.Ht.veg>                
>            =======

## C2 - Test Full Model
```{r,echo=F}
H.full.elev<-lm(ElevDif~
                    Myc.Frac+SRL+Max.Root.Loca+Log.Ht.veg+
                     Log.Lamina.thck+LDMC+Log.Leaf.Area+Leaf.Mass.Frac,
                 data=dat.8t)
summary(H.full.elev)$coefficients
AICc(H.full.elev)
```


```{r,echo=T}
AICc(H.full.elev)
```

>         =======
>         C2 - SUMMARY
>         Leaf Mass Fraction and Maximum Root Location are significant. 
>         Full model explains 33% of variance (Adjusted R2) with p=0.01
>         =======

## C3 - Find best subset of models with AICc dredge ####

```{r,echo=F}
   # Include:
   #        - all 1st order terms 
   #        - 2-way and 3-way trait interactions from regression tree
   #        - Don't include trait interaction from interaction screening.
         
          # rewrite H1 with na.action='fail' - required by dredge ()
          #
dat.8t.nona<-dat.8t[!is.na(dat.8t$ElevDif),]          
H.global.elev<-lm(ElevDif~ Log.Ht.veg+Max.Root.Loca+Log.Lamina.thck+LDMC+Log.Leaf.Area+Leaf.Mass.Frac+SRL+Myc.Frac+
             Leaf.Mass.Frac:Max.Root.Loca + Max.Root.Loca:Log.Ht.veg +
                Leaf.Mass.Frac:Max.Root.Loca:Log.Ht.veg,
       data=dat.8t.nona,na.action='na.fail')
```
```{r,echo=T,eval=T,message=F}
dredge.elev<-dredge(H.global.elev,beta='sd',rank="AICc",m.lim = c(0,6),extra=c("R^2"),trace=F)
dredge.elev[1:8]

```

>         =======
>         C3 - SUMMARY
>         Leaf.Mass.fraction and Max.Root.Location matter.  
>         Traits explains ~0.4 to 0.5 of variation in change in elevation
>         Best models include 3-5 traits
>         =======
